---
title: "Directed Study Data Cleaning and Regression Analysis"
editor: visual
author: "Jack Katz"
about: 
  template: trestles
---
# Have the Returns to English Proficiency Decreased for Immigrants to the United States?

Below is the abstract and r code for current economic research I am conducting on returns to English proficiency for Spanish-speaking immigrants and determinants of English proficiency for Spanish speaking immigrants.The code starts with retrieving the data from Integrated Public Use Microdata Series (IPUMS.org), cleaning the data, creating new variables, performing econometric analysis and using stargazer to create charts to dispaly the findings. The current iteration of the paper can be found on the home page under "Economic Research". I presented these findings at the March 2024 Eastern Economic Conference in Boston. 

## Abstract 

New immigrants need to be successfully economically and socially integrated if they are to bring
positive economic effects. One important piece of integration into a new society is language
proficiency. As the U.S. economy increasingly relies on immigrant labor, it will become even
more important to understand the returns to speaking English as an immigrant as well as and the
determinants of English proficiency among immigrants. I focus specifically on the return to
learning English as a Spanish speaker. As the U.S. becomes a more Spanish-speaker accessible
country, as more government and other agencies are offering forms in Spanish, as well as
broader use of Spanish and Spanish media in American culture, the incentive to learn English
may be decreasing for Spanish immigrants. \

This paper hypothesizes that the return to learning English has decreased over the years, both
because the United States has become a more Spanish-friendly country and because the number
of Spanish speakers has increased in the country over time. I analyze data from the American
Community Survey (ACS) from 1980-2019. I first consider the determinants of English
proficiency. I then use an instrumental variables approach to consider the effect of English
proficiency on income, the return to proficiency.\

While I find no significant change in returns to English proficiency from decade to decade, there
is a significant decrease between the 1980 sample and the 2019 sample. The results also show a
significant reduction in the income gap between female and male Hispanic immigrants. 

## Loading in Data

```{r}
library(ipumsr)
ddi <- read_ipums_ddi("usa_00018.xml")
data <- read_ipums_micro(ddi)
```


## Creating New Variables

In this section I recode the English proficiency variable, create a categorical 
education variable, create a variable to measure the age at arrival for all immigrants,
and adjust the income variable for inflation and convert it to log values. 
```{r, eval=FALSE}

library(tidyverse)
# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))
```

## Creating Samples 


### 1980 Sample
```{r}

library(tidyverse)
# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))





# 1980 --------------------------------------------------------------------
library(tidyverse)
library(AER)

Cities_1980 <- data %>% 
  filter(YEAR == 1980) %>% 
  filter(AAA %in% c(0:18)) %>% 
  filter(YRIMMIG %in% c(1950:1964)) %>% 
  mutate(Hispanic_Origin = ifelse(HISPAN %in% c(1:4), 1, 0)) %>% 
  mutate(Middle = ifelse(EDUC_Grade == "Middle", 1, 0)) %>% 
  mutate(High_School = ifelse(EDUC_Grade == "High_School", 1, 0)) %>% 
  mutate(College = ifelse(EDUC_Grade == "College", 1, 0)) %>% 
  mutate(Graduate_Degree = ifelse(EDUC_Grade == "Graduate_Degree", 1, 0)) %>% 
  mutate(Gender = ifelse(SEX == 2 , 1 , 0)) %>% 
  filter(Hispanic_Origin == 1) %>% 
  mutate(Citizenship_Status = ifelse(CITIZEN == 2, 1, 0)) %>% 
  filter(Graduate_Degree != 1) %>% 
  filter(ELDCH != 99 |
           YNGCH!= 99) %>% 
  filter(STATEFIP %in% c(6, 12, 36, 48, 34) |
           COUNTYFIP %in% c(37, 86, 73, 81, 215, 59, 85, 19, 47, 17)) %>% 
  mutate(YSM = (AGE - AAA)) %>% 
  filter(STATEFIP != 0 |
           COUNTYFIP != 0) %>% 
  mutate(Experience_Squared = ifelse(EDUC_Grade == "Elementary", ((AGE - 5 - 5)^2)/100, 
                                     ifelse(EDUC_Grade == "Middle", ((AGE - 8 - 5)^2)/100, 
                                            ifelse(EDUC_Grade == "High_School", ((AGE - 12 - 5)^2)/100, 
                                                   ifelse(EDUC_Grade == "College", ((AGE - 16 - 5)^2)/100, 0))))) %>% 
  mutate(Experience = ifelse(EDUC_Grade == "Elementary", AGE - 5 - 5, 
                             ifelse(EDUC_Grade == "Middle", AGE - 8 - 5, 
                                    ifelse(EDUC_Grade == "High_School", AGE - 12 - 5, 
                                           ifelse(EDUC_Grade == "College", AGE - 16 - 5, 0))))) %>% 
  mutate(Years_Education = ifelse(EDUC_Grade == "Elementary", 5, 
                                  ifelse(EDUC_Grade == "Middle", 8,
                                         ifelse(EDUC_Grade == "High_School", 12,
                                                ifelse(EDUC_Grade == "College", 16, 0))))) %>% 
  mutate(Education_Dummy = ifelse(EDUC_Grade %in% c("Elementary", "Middle"), 0, 1)) %>% 
  filter(INCTOT != 9999999 | INCTOT != 9999998) %>% 
  filter(ELDCH != 99 |
           YNGCH != 99) %>% 
  mutate(Hispanic_Population = 
           if_else(COUNTYFIP == 37, 27.6, 
                   if_else(COUNTYFIP == 86, 35.7, 
                           if_else(COUNTYFIP == 73, 14.8,
                                   if_else(COUNTYFIP == 81, 13.9,
                                           if_else(COUNTYFIP == 215, 81.3, 
                                                   
                                                   if_else(COUNTYFIP == 59, 14.8,
                                                           if_else(COUNTYFIP == 85, 17.5,
                                                                   if_else(COUNTYFIP == 19, 29.3,
                                                                           if_else(COUNTYFIP == 57, 17.6,
                                                                                   if_else(COUNTYFIP == 17, 26.1, 0))))))))))) %>% 
  filter(Hispanic_Population != 0)


view(head(Cities_1980, 100))




# OLS Regression on Log Wages
OLS1980 <- lm(LogWages ~ AGE + Gender  +  SPEAKENG_new + Citizenship_Status + Experience_Squared , data = Cities_1980)
summary(OLS1980)

#OLS Regression on SPEAKEN_eng

ENG1980 <- lm(SPEAKENG_new ~ AAA  + ELDCH + YNGCH + 
                Hispanic_Population + Years_Education,
              data = Cities_1980)
summary(ENG1980)

English1980 <- predict(ENG1980)
English1980


# IV Regression 1980

iv1980 <- ivreg(LogWages ~  SPEAKENG_new + Experience_Squared  + Gender  |    Experience_Squared   +
                  Gender + AAA + Hispanic_Population +
                  ELDCH + YNGCH + Years_Education , data = Cities_1980)
summary(iv1980)

Education1980 <- lm(LogWages ~ Years_Education + Experience + Experience_Squared + AGE + Gender, data = Cities_1980)
summary(Education1980)


Used_variables <- Cities_1980 %>% 
  select(AGE, SPEAKENG_new, Experience_Squared, Experience, AAA, Years_Education)



cor(Used_variables, use="complete.obs", method="pearson")

?cor

class(Cities_1980$COUNTY_POPULATION)




#Creating Population Concentration variable

Counties1980 <- Cities_1980 %>%
  filter(COUNTYFIP != 0) %>% 
  group_by(STATEFIP, COUNTYFIP) %>% 
  summarise(n = n()) %>% 
  arrange(-n)

# County Codes
# 37 = Los Angeles, California, 27.6% 

# 86 = Dade/Miami, Florida, 35.7%

# 73 = San Diego, California, 14.8%   

# 81 = Queens, New York, 13.9%

# 215 = Hidalgo, Texas, 81.3%

# 59 = Orange County, California, 14.8%

# 85 = Santa Clara County, California, 17.5%

# 19 = fresno county, California, 29.3%

# 57 = kings County, New york, 17.6% 

# 17 = hudson County, New Jersey, 26.1%



mutate(Cities_1980, Hispanic_Population = 
         if_else(COUNTYFIP == 37, 27.6, 
                 if_else(COUNTYFIP == 86, 35.7, 
                         if_else(COUNTYFIP == 73, 14.8,
                                 if_else(COUNTYFIP == 81, 13.9,
                                         if_else(COUNTYFIP == 215, 81.3, 
                                                 
                                                 if_else(COUNTYFIP == 59, 14.8,
                                                         if_else(COUNTYFIP == 85, 17.5,
                                                                 if_else(COUNTYFIP == 19, 29.3,
                                                                         if_else(COUNTYFIP == 57, 17.6,
                                                                                 if_else(COUNTYFIP == 17, 26.1, 0)))))))))))


view(head(Cities_1980, 10))


ENG80 <- lm(SPEAKENG_new ~ AAA + Citizenship_Status + ELDCH + YNGCH + Hispanic_Population + Citizenship_Status + Years_Education,
            data = Cities_1980)
summary(ENG80)


```
### 1990 Sample 

```{r}
library(tidyverse)
library(AER)

# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))



# 1990 --------------------------------------------------------------------
library(tidyverse)
# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))


Cities_1990 <- Sample1990 <- data %>% 
  filter(YEAR == 1990) %>% 
  filter(AAA %in% c(0:18)) %>% 
  filter(YRIMMIG %in% c(1960:1974)) %>% 
  mutate(Hispanic_Origin = ifelse(HISPAN %in% c(1:4), 1, 0)) %>% 
  mutate(Middle = ifelse(EDUC_Grade == "Middle", 1, 0)) %>% 
  mutate(High_School = ifelse(EDUC_Grade == "High_School", 1, 0)) %>% 
  mutate(College = ifelse(EDUC_Grade == "College", 1, 0)) %>% 
  mutate(Graduate_Degree = ifelse(EDUC_Grade == "Graduate_Degree", 1, 0)) %>% 
  mutate(Gender = ifelse(SEX == 2 , 1 , 0)) %>% 
  filter(Hispanic_Origin == 1) %>% 
  mutate(Citizenship_Status = ifelse(CITIZEN == 2, 1, 0)) %>% 
  filter(Graduate_Degree != 1) %>% 
  filter(ELDCH != 99 |
           YNGCH!= 99) %>% 
  filter(STATEFIP %in% c(6, 12, 17, 36)|
           COUNTYFIP %in% c(37, 86, 31, 59, 5, 48)) %>% 
  mutate(YSM = (AGE - AAA)) %>% 
  filter(STATEFIP != 0 |
           COUNTYFIP != 0)%>% 
  mutate(Experience_Squared = ifelse(EDUC_Grade == "Elementary", ((AGE - 5 - 5)^2)/100, 
                                     ifelse(EDUC_Grade == "Middle", ((AGE - 8 - 5)^2)/100, 
                                            ifelse(EDUC_Grade == "High_School", ((AGE - 12 - 5)^2)/100, 
                                                   ifelse(EDUC_Grade == "College", ((AGE - 16 - 5)^2)/100, 0))))) %>% 
  mutate(Experience = ifelse(EDUC_Grade == "Elementary", AGE - 5 - 5, 
                             ifelse(EDUC_Grade == "Middle", AGE - 8 - 5, 
                                    ifelse(EDUC_Grade == "High_School", AGE - 12 - 5, 
                                           ifelse(EDUC_Grade == "College", AGE - 16 - 5, 0))))) %>% 
  mutate(Years_Education = ifelse(EDUC_Grade == "Elementary", 5, 
                                  ifelse(EDUC_Grade == "Middle", 8,
                                         ifelse(EDUC_Grade == "High_School", 12,
                                                ifelse(EDUC_Grade == "College", 16, 0))))) %>% 
  mutate(Education_Dummy = ifelse(EDUC_Grade %in% c("Elementary", "Middle"), 0, 1)) %>% 
  filter(INCTOT != 9999999 | INCTOT != 9999998) %>% 
  mutate( Hispanic_Population = 
            if_else(COUNTYFIP == 37, 37.8 , 
                    if_else(COUNTYFIP == 86, 49.2 , 
                            if_else(COUNTYFIP == 31, 13.6 ,
                                    if_else(COUNTYFIP == 59, 23.4 ,
                                            if_else(COUNTYFIP == 5, 43.5,
                                                    if_else(COUNTYFIP == 47, 20.1,
                                                            if_else(COUNTYFIP == 73, 20.4,
                                                                    if_else(COUNTYFIP == 81, 19.5,
                                                                            if_else(COUNTYFIP == 61, 26.0,
                                                                                    if_else(COUNTYFIP == 141, 69.6, 0))))))))))) %>% 
  filter(Hispanic_Population != 0)




#1990 OLS 
OLS1990 <- lm(LogWages ~ AGE + Gender  +  SPEAKENG_new + Citizenship_Status + Experience_Squared, data = Cities_1990)
summary(OLS1990)

#1990 OLS on SPEAKENG
ENG1990 <- lm(SPEAKENG_new ~ AAA + ELDCH + YNGCH + 
                Hispanic_Population + Years_Education,
              data = Cities_1990)
summary(ENG1990)

#1990 IV 
iv1990 <- ivreg(LogWages ~  SPEAKENG_new + Experience_Squared + Gender|Experience_Squared   +
                  Gender + AAA + Hispanic_Population +
                  ELDCH + YNGCH + Years_Education , data = Cities_1990)

summary(iv1990)


#Creating Population variable 

Counties_1990 <- Cities_1990 %>% 
  filter(COUNTYFIP != 0) %>% 
  group_by(STATEFIP, COUNTYFIP) %>% 
  summarise(n=n()) %>% 
  arrange(-n)

# County codes

# 37 = Los Angeles, California, 37.8

# 86 = Miami/Dade county, 49.2

# 31 = Cook County, Illinois

# 59 = Orange County, California 

# 5 = Bronx County, New York


Cities_1990 <- mutate(Cities_1990, Hispanic_Population = 
                        if_else(COUNTYFIP == 37, 37.8 , 
                                if_else(COUNTYFIP == 86, 49.2 , 
                                        if_else(COUNTYFIP == 31, 13.6 ,
                                                if_else(COUNTYFIP == 59, 23.4 ,
                                                        if_else(COUNTYFIP == 5, 43.5,
                                                                if_else(COUNTYFIP == 47, 20.1,
                                                                        if_else(COUNTYFIP == 73, 20.4,
                                                                                if_else(COUNTYFIP == 81, 19.5,
                                                                                        if_else(COUNTYFIP == 61, 26.0,
                                                                                                if_else(COUNTYFIP == 141, 69.6, 0)))))))))))

view(head(Cities_1990))


```
### 2000 Sample

```{r}
library(tidyverse)
library(AER)

# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))



# 2000 --------------------------------------------------------------------

library(tidyverse)
# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))



Cities_2000 <- data %>% 
  filter(YEAR == 2000) %>% 
  filter(AAA %in% c(0:18)) %>% 
  filter(YRIMMIG %in% c(1970:1984)) %>% 
  mutate(Hispanic_Origin = ifelse(HISPAN %in% c(1:4), 1, 0)) %>% 
  mutate(Middle = ifelse(EDUC_Grade == "Middle", 1, 0)) %>% 
  mutate(High_School = ifelse(EDUC_Grade == "High_School", 1, 0)) %>% 
  mutate(College = ifelse(EDUC_Grade == "College", 1, 0)) %>% 
  mutate(Graduate_Degree = ifelse(EDUC_Grade == "Graduate_Degree", 1, 0)) %>% 
  mutate(Gender = ifelse(SEX == 2 , 1 , 0)) %>% 
  filter(Hispanic_Origin == 1) %>% 
  mutate(Citizenship_Status = ifelse(CITIZEN == 2, 1, 0)) %>% 
  filter(Graduate_Degree != 1) %>% 
  filter(ELDCH != 99 |
           YNGCH!= 99) %>% 
  filter(STATEFIP %in% c(6, 48, 17, 4, 36) |
           COUNTYFIP %in% c(37, 59, 201, 31, 73, 71, 113, 13, 81, 5)) %>% 
  mutate(YSM = (AGE - AAA)) %>% 
  filter(STATEFIP != 0 |
           COUNTYFIP != 0) %>% 
  mutate(Experience = ifelse(EDUC_Grade == "Elementary", AGE - 5 - 5, 
                             ifelse(EDUC_Grade == "Middle", AGE - 8 - 5, 
                                    ifelse(EDUC_Grade == "High_School", AGE - 12 - 5, 
                                           ifelse(EDUC_Grade == "College", AGE - 16 - 5, 0))))) %>% 
  mutate(Experience_Squared = ifelse(EDUC_Grade == "Elementary", ((AGE - 5 - 5)^2)/100, 
                                     ifelse(EDUC_Grade == "Middle", ((AGE - 8 - 5)^2)/100, 
                                            ifelse(EDUC_Grade == "High_School", ((AGE - 12 - 5)^2)/100, 
                                                   ifelse(EDUC_Grade == "College", ((AGE - 16 - 5)^2)/100, 0))))) %>% 
  mutate(Years_Education = ifelse(EDUC_Grade == "Elementary", 5, 
                                  ifelse(EDUC_Grade == "Middle", 8,
                                         ifelse(EDUC_Grade == "High_School", 12,
                                                ifelse(EDUC_Grade == "College", 16, 0)))))%>% 
  mutate(Education_Dummy = ifelse(EDUC_Grade %in% c("Elementary", "Middle"), 0, 1)) %>% 
  filter(INCTOT != 9999999 | INCTOT != 9999998) %>% 
  mutate(Hispanic_Population = 
           if_else(COUNTYFIP == 37, 44.6, 
                   if_else(COUNTYFIP == 59, 30.8, 
                           if_else(COUNTYFIP == 201, 32.9,
                                   if_else(COUNTYFIP == 31, 19.9,
                                           if_else(COUNTYFIP == 73, 26.7, 
                                                   if_else(COUNTYFIP == 71,39.2 , 
                                                           if_else(COUNTYFIP == 113,29.9 , 
                                                                   if_else(COUNTYFIP == 13, 24.8, 
                                                                           if_else(COUNTYFIP == 81,25.0 , 
                                                                                   if_else(COUNTYFIP == 5,48.4 ,0))))))))))) %>% 
  filter(Hispanic_Population != 0)



#  filter(COUNTYFIP %in% c(37, 59, 31, 201, 73))


#Creating Population variable
Counties2000 <- Cities_2000 %>% 
  filter(COUNTYFIP != 0 ) %>% 
  group_by(STATEFIP, COUNTYFIP) %>% 
  summarise(n=n()) %>% 
  arrange(-n)

# 37 = Los Angeles, 44.6

# 59 = Daviess County, Kentucky, 0.9

# 31 = Whitefeld, Georgia, 22.1

# 201 = Winnebago, Illinois, 6.9

# Colombia, Georgia, 2.6

Cities_2000 <- mutate(Cities_2000, Hispanic_Population = 
                        if_else(COUNTYFIP == 37, 44.6, 
                                if_else(COUNTYFIP == 59, 30.8, 
                                        if_else(COUNTYFIP == 201, 32.9,
                                                if_else(COUNTYFIP == 31, 19.9,
                                                        if_else(COUNTYFIP == 73, 26.7, 
                                                                if_else(COUNTYFIP == 71,39.2 , 
                                                                        if_else(COUNTYFIP == 113,29.9 , 
                                                                                if_else(COUNTYFIP == 13, 24.8, 
                                                                                        if_else(COUNTYFIP == 81,25.0 , 
                                                                                                if_else(COUNTYFIP == 5,48.4 ,0)))))))))))                  




#OLS 2000
OLS2000 <- lm(LogWages ~ AGE + Gender  + SPEAKENG_new + Citizenship_Status + Experience_Squared, data = Cities_2000)
summary(OLS2000)

#OlS on SPEAKEng

ENG2000 <- lm(SPEAKENG_new ~ AAA  + ELDCH + YNGCH + 
                Hispanic_Population + Years_Education
              ,
              data = Cities_2000)
summary(ENG2000)

#2000 IV 
iv2000 <- ivreg(LogWages ~  SPEAKENG_new + Experience_Squared   + Gender|    Experience_Squared  + 
                  Gender + AAA + Hispanic_Population +
                  ELDCH + YNGCH + Years_Education , data = Cities_2000)
summary(iv2000)







```
### 2010 Sample 
```{r}

# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))



# 2010 --------------------------------------------------------------------
library(tidyverse)
# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))


library(tidyverse)
library(AER)

Cities_2010 <- data %>% 
  filter(YEAR == 2010) %>% 
  filter(AAA %in% c(0:18)) %>% 
  filter(YRIMMIG %in% c(1980:1994)) %>% 
  mutate(Hispanic_Origin = ifelse(HISPAN %in% c(1:4), 1, 0)) %>% 
  mutate(Middle = ifelse(EDUC_Grade == "Middle", 1, 0)) %>% 
  mutate(High_School = ifelse(EDUC_Grade == "High_School", 1, 0)) %>% 
  mutate(College = ifelse(EDUC_Grade == "College", 1, 0)) %>% 
  mutate(Graduate_Degree = ifelse(EDUC_Grade == "Graduate_Degree", 1, 0)) %>% 
  mutate(Gender = ifelse(SEX == 2 , 1 , 0)) %>% 
  filter(Hispanic_Origin == 1) %>% 
  mutate(Citizenship_Status = ifelse(CITIZEN == 2, 1, 0)) %>% 
  filter(Graduate_Degree != 1) %>% 
  filter(ELDCH != 99 |
           YNGCH!= 99) %>% 
  filter(COUNTYFIP %in% c(37, 201, 31, 59, 13)) %>% 
  filter(STATEFIP %in% c(6, 48, 17, 4, 36, 24, 34,32) |
           COUNTYFIP %in% c(37, 201, 59, 31, 13)) %>% 
  mutate(YSM = (AGE - AAA)) %>% 
  filter(STATEFIP != 0 |
           COUNTYFIP != 0) %>% 
  mutate(Experience = ifelse(EDUC_Grade == "Elementary", AGE - 5 - 5, 
                             ifelse(EDUC_Grade == "Middle", AGE - 8 - 5, 
                                    ifelse(EDUC_Grade == "High_School", AGE - 12 - 5, 
                                           ifelse(EDUC_Grade == "College", AGE - 16 - 5, 0))))) %>% 
  mutate(Experience_Squared = ifelse(EDUC_Grade == "Elementary", ((AGE - 5 - 5)^2)/100, 
                                     ifelse(EDUC_Grade == "Middle", ((AGE - 8 - 5)^2)/100, 
                                            ifelse(EDUC_Grade == "High_School", ((AGE - 12 - 5)^2)/100, 
                                                   ifelse(EDUC_Grade == "College", ((AGE - 16 - 5)^2)/100, 0))))) %>% 
  mutate(Years_Education = ifelse(EDUC_Grade == "Elementary", 5, 
                                  ifelse(EDUC_Grade == "Middle", 8,
                                         ifelse(EDUC_Grade == "High_School", 12,
                                                ifelse(EDUC_Grade == "College", 16, 0)))))%>% 
  mutate(Education_Dummy = ifelse(EDUC_Grade %in% c("Elementary", "Middle"), 0, 1)) %>% 
  filter(INCTOT != 9999999 | INCTOT != 9999998) %>% 
  mutate(Hispanic_Population = 
           if_else(COUNTYFIP == 37,47.7 , 
                   if_else(COUNTYFIP == 201, 11.1, 
                           if_else(COUNTYFIP == 59 | STATEFIP == 6, 33.7,
                                   if_else(COUNTYFIP == 31, 24.0 ,
                                           if_else(COUNTYFIP == 13 | STATEFIP == 4,29.6 , 
                                                   if_else(COUNTYFIP == 13 | STATEFIP == 6, 24.4,
                                                           if_else(COUNTYFIP == 59 | STATEFIP == 36, 14.6 ,
                                                                   if_else(COUNTYFIP == 31 | STATEFIP == 24, 17.0,
                                                                           if_else(COUNTYFIP == 13 | STATEFIP == 34,20.3 , 
                                                                                   if_else(COUNTYFIP == 31 | STATEFIP == 32,22.2 ,0))))))))))) %>% 
  filter(Hispanic_Population != 0)

#Creating population variable 

Cities_2010 %>% 
  filter(COUNTYFIP != 0 ) %>% 
  group_by(STATEFIP, COUNTYFIP) %>% 
  summarise(n=n()) %>% 
  arrange(-n)

# 37 = Los Angeles, California, 47.7

# 201 = Winnebago, Illinois, 10.9

# 31 = Witefeld, Georgia, 31.6

# 59 = Daviess County, Kentucky, 2.6 

# 13 = Black Hawk Iowa, 3.7

Cities_2010 <- mutate(Cities_2010, Hispanic_Population = 
                        if_else(COUNTYFIP == 37,47.7 , 
                                if_else(COUNTYFIP == 201, 11.1, 
                                        if_else(COUNTYFIP == 59 | STATEFIP == 6, 33.7,
                                                if_else(COUNTYFIP == 31, 24.0 ,
                                                        if_else(COUNTYFIP == 13 | STATEFIP == 4,29.6 , 
                                                                if_else(COUNTYFIP == 13 | STATEFIP == 6, 24.4,
                                                                        if_else(COUNTYFIP == 59 | STATEFIP == 36, 14.6 ,
                                                                                if_else(COUNTYFIP == 31 | STATEFIP == 24, 17.0,
                                                                                        if_else(COUNTYFIP == 13 | STATEFIP == 34,20.3 , 
                                                                                                if_else(COUNTYFIP == 31 | STATEFIP == 32,22.2 ,0)))))))))))

#OLS 2010
OLS2010 <- lm(LogWages ~ AGE + Gender  +   SPEAKENG_new  + Citizenship_Status + Experience_Squared, data = Cities_2010)
summary(OLS2010)

#OlS on SPEAKEng
ENG2010 <- lm(SPEAKENG_new ~ AAA  + ELDCH + YNGCH + 
                Hispanic_Population + Years_Education 
              ,
              data = Cities_2010)
summary(ENG2010)

#IV 2010
iv2010 <- ivreg(LogWages ~  SPEAKENG_new + Experience_Squared  + Gender |    Experience_Squared  + 
                  Gender + AAA + Hispanic_Population +
                  ELDCH + YNGCH + Years_Education, data = Cities_2010)
summary(iv2010)

view(head(Cities_1980, 10))


```
### 2019 Sample
```{r}

# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))



# 2019 --------------------------------------------------------------------
library(tidyverse)
# SPEAKENG

data$SPEAKENG_numeric <- as.numeric(data$SPEAKENG)
data$SPEAKENG_new <- dplyr::recode(data$SPEAKENG_numeric, `3` = 1, `4` = 2, `5` = 3, `6` = 4, `1` = 5)

#EDUC

data$EDUC_Grade <- case_when(data$EDUC %in% c(0:1) ~ "Elementary",
                             data$EDUC %in% c(2) ~ "Middle",
                             data$EDUC %in% c(3:6) ~ "High_School",
                             data$EDUC %in% c(7:10) ~ "College",
                             data$EDUC %in% c(11) ~ "Graduate_Degree")

#Age At Arrival Variable 

data <- mutate(data, AAA = AGE - (YEAR - YRIMMIG))

#Inflation Adjusted Income Variable 

data <- filter(data, INCTOT > 0) 
data <- mutate(data, Inflation_Adjusted = INCTOT*CPI99)

#Log wages
data <- mutate(data, LogWages = log(Inflation_Adjusted))


library(tidyverse)
library(AER)

Cities_2019 <- data %>% 
  filter(YEAR == 2019) %>% 
  filter(AAA %in% c(0:18)) %>% 
  filter(YRIMMIG %in% c(1989:2003)) %>% 
  mutate(Hispanic_Origin = ifelse(HISPAN %in% c(1:4), 1, 0)) %>% 
  mutate(Middle = ifelse(EDUC_Grade == "Middle", 1, 0)) %>% 
  mutate(High_School = ifelse(EDUC_Grade == "High_School", 1, 0)) %>% 
  mutate(College = ifelse(EDUC_Grade == "College", 1, 0)) %>% 
  mutate(Graduate_Degree = ifelse(EDUC_Grade == "Graduate_Degree", 1, 0)) %>% 
  mutate(Gender = ifelse(SEX == 2 , 1 , 0)) %>% 
  filter(Hispanic_Origin == 1) %>% 
  mutate(Citizenship_Status = ifelse(CITIZEN == 2, 1, 0)) %>%
  filter(Graduate_Degree != 1) %>% 
  filter(ELDCH != 99 |
           YNGCH != 99) %>% 
  filter(COUNTYFIP %in% c(37, 31, 201, 113, 13)) %>% 
  mutate(Years_In_Country = (AGE- AAA)) %>% 
  filter(STATEFIP %in% c(6, 48, 17, 4, 24, 32, 34) |
           COUNTYFIP %in% c(37, 201, 113, 31, 13)) %>% 
  mutate(YSM = (AGE - AAA)) %>% 
  filter(STATEFIP != 0 |
           COUNTYFIP != 0) %>% 
  mutate(Experience = ifelse(EDUC_Grade == "Elementary", AGE - 5 - 5, 
                             ifelse(EDUC_Grade == "Middle", AGE - 8 - 5, 
                                    ifelse(EDUC_Grade == "High_School", AGE - 12 - 5, 
                                           ifelse(EDUC_Grade == "College", AGE - 16 - 5, 0))))) %>% 
  mutate(Experience_Squared = ifelse(EDUC_Grade == "Elementary", ((AGE - 5 - 5)^2)/100, 
                                     ifelse(EDUC_Grade == "Middle", ((AGE - 8 - 5)^2)/100, 
                                            ifelse(EDUC_Grade == "High_School", ((AGE - 12 - 5)^2)/100, 
                                                   ifelse(EDUC_Grade == "College", ((AGE - 16 - 5)^2)/100, 0))))) %>% 
  mutate(Education_Dummy = ifelse(EDUC_Grade %in% c("Elementary", "Middle"), 0, 1)) %>% 
  mutate(Years_Education = ifelse(EDUC_Grade == "Elementary", 5, 
                                  ifelse(EDUC_Grade == "Middle", 8,
                                         ifelse(EDUC_Grade == "High_School", 12,
                                                ifelse(EDUC_Grade == "College", 16, 0))))) %>% 
  filter(INCTOT != 9999999 | INCTOT != 9999998) %>% 
  mutate(Hispanic_Population = 
           ifelse(COUNTYFIP == 37 | STATEFIP == 6, 48.0,
                  ifelse(COUNTYFIP == 201 | STATEFIP == 48, 43.0,
                         ifelse(COUNTYFIP == 113 | STATEFIP == 48, 40.5,
                                ifelse(COUNTYFIP == 31 | STATEFIP == 17, 26.2,
                                       ifelse(COUNTYFIP == 13 | STATEFIP == 4, 30.6,
                                              ifelse(COUNTYFIP == 13 | STATEFIP == 6, 27.0,
                                                     ifelse(COUNTYFIP == 31 | STATEFIP == 24, 20.5,
                                                            ifelse(COUNTYFIP == 31 | STATEFIP == 34, 42.7, 
                                                                   ifelse(COUNTYFIP == 31 | STATEFIP == 32, 25.1,
                                                                          ifelse(COUNTYFIP == 13 | STATEFIP == 34, 24.4, 0))))))))))) %>% 
  filter(Hispanic_Population != 0 )


#mutate(COUNTY_POPULATION = ifelse(STATEFIP %in% c(6,12, 36, 48, 36, 34), 1, 0))

view(head(Cities_2019
))

# Creating Population Variable 
Cities_2019 %>% 
  group_by(STATEFIP, COUNTYFIP) %>% 
  summarise(n=n()) %>% 
  arrange(-n)

Cities_2019 <- mutate(Cities_2019, Hispanic_Population = 
                        ifelse(COUNTYFIP == 37 | STATEFIP == 6, 48.0,
                               ifelse(COUNTYFIP == 201 | STATEFIP == 48, 43.0,
                                      ifelse(COUNTYFIP == 113 | STATEFIP == 48, 40.5,
                                             ifelse(COUNTYFIP == 31 | STATEFIP == 17, 26.2,
                                                    ifelse(COUNTYFIP == 13 | STATEFIP == 4, 30.6,
                                                           ifelse(COUNTYFIP == 13 | STATEFIP == 6, 27.0,
                                                                  ifelse(COUNTYFIP == 31 | STATEFIP == 24, 20.5,
                                                                         ifelse(COUNTYFIP == 31 | STATEFIP == 34, 42.7, 
                                                                                ifelse(COUNTYFIP == 31 | STATEFIP == 32, 25.1,
                                                                                       ifelse(COUNTYFIP == 13 | STATEFIP == 34, 24.4, 0)))))))))))

#OLS 2019
OLS2019 <- lm(LogWages ~ AGE + Gender   +SPEAKENG_new + Citizenship_Status + Experience_Squared  , data = Cities_2019)
summary(OLS2019)


#OlS on SPEAKEng
ENG2019 <- lm(SPEAKENG_new ~ AAA  + ELDCH + YNGCH + 
                Hispanic_Population + Years_Education 
              ,
              data = Cities_2019)
summary(ENG2019)

#IV 2019
iv2019 <- ivreg(LogWages ~  SPEAKENG_new + Experience_Squared  + Gender|    Experience_Squared  + 
                  Gender + AAA + Hispanic_Population +
                  ELDCH + YNGCH  + Years_Education, data = Cities_2019)
summary(iv2019)

```

## Regression Results 
```{r, eval=FALSE}
# 1980 

# Multiple regression
ENG1980 <- lm(SPEAKENG_new ~ AAA  + ELDCH + YNGCH + 
                Hispanic_Population + Years_Education,
              data = Cities_1980)
summary(ENG1980)

```